---
---
title: "Modeling and prediction for movies"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup
install.packages("rcompanion")
### Load packages
```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `movies`. Delete this note when before you submit 
your work. 

```{r load-data}
load(file.choose())
```



* * *

## Part 1: Data
<ol>
The data is a recollection from 651 random sampled movies that were filmed and released before 2016. This limits the scope of the research to an observational study were no causality can be inferred. Also the results may not be generalizable to audiences beyond those that visit pages like Rotten Tomatoes (critics and audience). Further, generalizability also gets hurt from the fact that may only include movies that are suitable for western societies.
</ol>
* * *

## Part 2: Research question
<ol>
Is there a relationship between the rating at audience score on Rotten Tomatoes (audience_score) the genre (genre), the duration of the movie (runtime), critics score (critics_score), and the rating of the movie (mpaa_rating).

In a nutshell, I would like to know if there is a relation between the above mentioned variables and the score in Rotten Tomatoes of the audience. 

These questions are interesting since it would help use to make a sort of formula for getting the audience to the sack; for predicting the type of movies they tend to love and dislike.

</ol>
* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button above. Make sure that your code is visible in the project you submit. 
Delete this note when before you submit your work.

* * *
<ol>
First, we need to see the type of data we are working with
</ol>
```{r data str}
str(movies)
```
<ol>
Therefore our variables of interest are:
1)runtime is numeric

2)genre with 11 levels (Action & Adventure, Comedy, Documentary, Drama, Horror, Mystery & Suspense, Other)

3)mpaa_rating with 6 levels

4)audience_score is numeric

5)critics_score is numeric

 With this in mind we now can begin to graphic our variables
</ol>
```{r Relations_audience}
#Genre
ggplot(data = movies, aes(x=genre, y=audience_score))+
  geom_boxplot(fill="transparent", color= "red", font= 1)+
  theme(text = element_text(size = 5))

```
<ol>
With this we can appreciate that there are some genres that have better scores like: documentary or musical. There is also others that seem to be very volatile like science fiction were there are some very low and high scores. These scores seem to suggest that there are some genres that are a safer bet than others.

Now lets see if there is some relation between audience_score and thtr_rel_month
</ol>
```{r audience_score and release_date}
ggplot(data = movies, aes(x=thtr_rel_month, y=audience_score))+
  geom_jitter(color= "light blue")+
   geom_smooth(method = "lm")
```
<ol>
Unfortunately there seem to be no linear relation between thtr_rel_month and audience_score. Therefore is not a good variable for the model

Now lets see how does runtime and audience_score.
</ol>
```{r runtime and score}
ggplot(data = movies, aes(x=runtime, y=audience_score))+
  geom_jitter(color="light blue")+
  geom_smooth(method = "lm")
```
<ol>
Again it seem that this are not good data since there are some extreme outliers that seem to affect the slope of the regression line. Still later analyses would tell.

Now lets check the relation between mpaa_rating and the audience_score
</ol>
```{r rating and score}
ggplot(data=movies, aes(x=mpaa_rating, y=audience_score))+
  geom_boxplot(color= "red")
```
<ol>
We can appreciate that the NC-17 (Adults Only) seem to be the less variable but also the ones with the least higher scores. This may because less people watch this movies. Also we can appreciate that the G (general audience movies) and the Unrated movies seem to be the ones with the better scores.

Finally, the relation between audience score and the critics score
</ol>
```{r audience and critics score}
ggplot(data=movies, aes(x=critics_score, y=audience_score))+
  geom_point(fill="light blue", size=2)+
  geom_smooth(method = "lm")
```
<ol>
Here we can appreciate 

## Part 4: Modeling

Now lets begin with our model
```{r  linear model}
m1 <- lm(audience_score~ genre+ critics_score+ mpaa_rating, runtime, data=movies)
summary(m1)
```
<ol>
Let's make some diagnostics
</ol>
```{r}
#Without genre
m2 <- lm(audience_score~critics_score+ mpaa_rating +runtime, data=movies)
summary(m2)$adj.r.squared
```
```{r}
#Without critics score
m2 <- lm(audience_score~genre+ mpaa_rating+runtime, data=movies)
summary(m2)$adj.r.squared
```
```{r}
#Without mpaa
m2 <- lm(audience_score~genre+ critics_score+runtime, data=movies)
summary(m2)$adj.r.squared
```

```{r}
#Without runtime
m2 <- lm(audience_score~genre+ critics_score+ mpaa_rating, data=movies)
summary(m2)$adj.r.squared 
```
<ol>
Because of this we stay with m1 since the R squared is bigger. This means, simply, that m1 explains more of the observed variance than m2.

Our linear model has the form: 
\[
audience\_score =   17.56732  + \hat{\beta}_{1j} \times genre_j + \hat{\beta}_{2} \times critics\_score + \hat{\beta}_{3j} \times mpaa\_rating_j +  \hat{\beta}_{4} \times  runtime
\]

In this model our coefficients mean that if X coefficient is the case then audience_score increases(or reduces) by X coefficient if, and only if, everything remains equal. Therefore coefficients effects are dependent on other coefficients for their particular relation to the outcome variable. Which is different from our normal linear regression where the relation is direct and independent.

Now let's make some diagnostics for our m1
</ol>

```{r diagnostics}
#homeoscedasticity
ggplot(data=m1, aes(x=.fitted, y=.resid))+
  geom_jitter(size= 2)+
  geom_hline(yintercept = 0, linetype= "dashed")+
  xlab("Fitted")+
  ylab("Residuals")

#Residuals normality

ggplot(data = m1, aes(x=.resid))+
  geom_histogram(binwidth = 5, fill= "light blue", color= "black")

```
<ol>
It seems that our model barely fits the conditions of homoscedasticity but definitely fulfills the normality of residuals, and as we saw above also de linearity conditions for runtime. Therefore we can continue to the prediction section.

</ol>

## Part 5: Prediction

<ol> 
For this lets suppose we have an hypothetical movie with a genre of "horror", a "Nc-17" grading, a critics score of 90, and with a runtime of 129 minutes.

Let's be sure that this are legit values and that we are not extrapolating our data

</ol>
```{r extrapolation or not}
summary(movies$critics_score)
summary(movies$runtime)

ggplot(data=movies, aes(x=genre, y=runtime))+
  geom_boxplot()+
  theme(text = element_text(size = 5))

ggplot(data = movies, aes(x=genre, y=critics_score))+
 geom_boxplot()+
  theme(text = element_text(size = 5))

ggplot(data = movies, aes(x=genre, y=mpaa_rating))+
  geom_count(fill= "light blue")+
  theme(text=element_text(size = 5))
  
```
<ol>
It seems that our numeric variables are a little bit outside our data, so let's procede to select some more adequate values like:
"horror", a "R" grading, a critics score of 70, and with a runtime of 100 minutes.

</ol>
```{r}
newmovie <- data.frame(genre = "Horror", critics_score = 70, runtime= 100, mpaa_rating= "R")

predict(m1, newmovie, interval = "prediction", level = 0.95)
```
<ol>
As we can see our model predicts a fitted audince_score of 40.5529 when we have such a such predictors, and with 95% of confidence that our audience_score would be between (11.63625, 67.27433).
</ol>

* * *

## Part 6: Conclusion

<ol>

As we saw our model does not have the most accurate prediction one can ask for. This may be due to the lack of data in horror movies overall and to the overall variability of the data. Nontheless it is a kind of a good estimate for movie makers and producers since they can use this model to predict that horror movies are not the safest bet for winning the audiences. As we saw before, probably art movies and documentaries are a safer bet for their money because their scores are less variable and tend to be on the higher end!.

</ol>
